Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-11 write-ups/PLDI/paper.tex

\paragraph{Work stealing example}  \ADComment{Tyler/Hugues: can you please say in the first paragraph that this has been implemented on GPUs and cite the right papers?}
%
The work stealing paradigm is useful when a computation can be split
into multiple tasks that can be processed independently, but where the
number of tasks is not known ahead of time because processing a given
task may create an arbitrary number of new tasks. Work stealing enables
dynamic balancing of tasks across several processing units, OpenCL
workgroups in our case. Each workgroup has an associated task pool from
which it obtains tasks to process, and to which it stores newly created
tasks. When a workgroup's pool becomes empty, a workgroup---rather than
terminating---tries to \emph{steal} a task from an other
pool. Workgroups keep on processing tasks until all the work is done.

@@

Our $\globalbarrier$ primitive is a kernel-level function: if it
appears in conditional code then it must be reached by \emph{all}
threads executing the cooperative kernel.  On reaching a
$\globalbarrier()$, a thread waits until all threads have arrived at
the barrier.  Once all threads have arrived, the threads may proceed
past the barrier with the guarantee that all global memory accesses
issued before the barrier have completed.  The $\globalbarrier$
primitive can be implemented by adapting an inter-workgroup barrier
design, e.g.~\cite{XF}, to take account of a growing and shrinking number of workgroups, and the atomic operations provided by
the OpenCL 2.0 memory model enable a memory-safe
implementation~\cite{DBLP:conf/oopsla/SorensenDBGR16}.  However, implementing such a barrier is
involved, hence why we include this function as a primitive.

>>>>

\paragraph{Work stealing example}
%
The work stealing paradigm is useful when a computation can be split
into multiple tasks that can be processed independently, but where the
number of tasks is not known ahead of time because processing a given
task may create an arbitrary number of new tasks. Work stealing enables
dynamic balancing of tasks across several processing units, OpenCL
workgroups in our case. Each workgroup has an associated task pool from
which it obtains tasks to process, and to which it stores newly created
tasks. When a workgroup's pool becomes empty, a workgroup---rather than
terminating---tries to \emph{steal} a task from an other
pool. Workgroups keep on processing tasks until all the work is
done. Previous work addresses the feasibility and efficiency of work
stealing on GPUs~ref\cite{dlb-web,TPO10}.

@@

Our $\globalbarrier$ primitive is a kernel-level function: if it
appears in conditional code then it must be reached by \emph{all}
threads executing the cooperative kernel.  On reaching a
$\globalbarrier()$, a thread waits until all threads have arrived at
the barrier.  Once all threads have arrived, the threads may proceed
past the barrier with the guarantee that all global memory accesses
issued before the barrier have completed.  The $\globalbarrier$
primitive can be implemented by adapting an inter-workgroup barrier
design, e.g.~\cite{XF10}, to take account of a growing and shrinking number of workgroups, and the atomic operations provided by
the OpenCL 2.0 memory model enable a memory-safe
implementation~\cite{DBLP:conf/oopsla/SorensenDBGR16}.  However, implementing such a barrier is
involved, hence why we include this function as a primitive.
